### 2013-02-15 doc/icml2013/algo.tex
Minor fixes

%%% second moments
Intuitively, performing regression on only $y$ given $x$ provides only first-order
information.  The key insight is that we can perform regression
on higher order powers to obtain more information about the parameters.
Specifically, for an integer $p \ge 1$, let us define the average
$m$-th order tensor power of the parameters as follows:
\begin{align}
M_p &\eqdef \sum_{h=1}^k \pi_h \beta_h\tp{p}. \label{eqn:Mp} % \in (\Re^{d})\tp{p}.
\end{align}
Now consider performing regression on $y^2$ given $x\tp{2}$.
Expanding $y^2 = (\innerp{\beta_h}{x} + \epsilon)^2$,
using the fact that $\innerp{\beta_h}{x}^p = \innerp{\beta_h\tp{p}}{x\tp{p}}$,
we have:
\begin{align}
y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + (\epsilon^2 - \sigma^2). \nonumber
\end{align}
Again, we have expressed $y^2$ has a linear function of $x\tp{2}$
with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.\footnote{If $\sigma^2$ were not known,
we could treat it as another coefficient
to be estimated.  The coefficients $M_2$ and $\sigma^2$ can be estimated jointly
provided that $x$ does not already contain a bias ($x_j$ must be non-constant for every $j \in [d]$).}
Importantly, the noise has mean zero; 
in fact each of the three terms has mean zero
by (i) definition of $M_2$, (ii) independence of $\epsilon$ and $h$,
and (iii) the fact that $\E[\epsilon^2] = \sigma^2$.

@@

%%% third moment
Let us look to the third moment for additional information.
We can write $y^3$ as a linear function of $x\tp{3}$ with coefficients $M_3$
plus a bias $3 \sigma^2 \innerp{M_1}{x}$ that can be estimated from \refeqn{y1} and zero mean noise $\eta_3(x)$:
\begin{align}
y^3 &= \innerp{M_3}{x\tp{3}} + 3\sigma^2 \innerp{M_1}{x} + \eta_3(x), \label{eqn:y3} \\
\eta_3(x) &= \innerp{\beta_h\tp{3} - M_3}{x\tp{3}}
+ 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \nonumber \\
&\quad + 3(\epsilon^2 \innerp{\beta_h}{x} - \sigma^2 \innerp{M_1}{x})
+ \epsilon^3. \nonumber
\end{align}
Performing regression yields estimates for $M_3$.
It turns out that knowledge of $M_2$ and $M_3$ are sufficient to recover
all the parameters.

% Full algorithm
Now we are ready to state our full algorithm, which we call Spectral Experts
(\algorithmref{sec:algo}).
First, we perform three regressions to recover the compound parameters
$M_1$ \refeqn{y1},
$M_2$ \refeqn{y2}, and
$M_3$ \refeqn{y3}.
Since $M_2$ and $M_3$ both only have rank $k$,
we can use nuclear norm regularization
\cite{Tomioka2011,NegahbanWainwright2009}
to exploit this low-rank structure and improve our compound parameter estimates.
In the algorithm, the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(2)}$
are set to $\frac{1}{\sqrt{n}}$.
The resulting semidefinite program is a standard one which has received
much attention in recent years.
We use a rather simple gradient-based approach,
where the nuclear norm is handled in closed form by taking an SVD
and soft-thresholding the eigenvalues \cite{donoho95soft,cai10soft}.

% Tensor factorization
Having estimated the compound parameters $M_2$ and $M_3$,
it remains to recover the original parameters $\theta$.
\citet{AnandkumarGeHsu2012} showed that for $M_2$ and $M_3$ of
the forms in \refeqn{Mp}, it is possible to efficiently accomplish this.
Specifically, we first compute a whitening matrix based on the SVD of $M_2$
and use that to construct a tensor $T = M_3(W, W, W)$ whose factors are orthogonal.
We can use the robust tensor power method to compute all the
eigenvalues and eigenvectors of $T$, from which it is easy to recover
the parameters $\pi$ and $\{\beta_h\}$.

@@

In this section, we conduct a theoretical analysis of Spectral Experts.
In particular, we prove that it enjoys consistent parameter estimates.
\begin{theorem}[Convergence of Spectral Experts]
  \label{thm:convergence}
Each $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Assume that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for each $p \in \{1,2,3\}$.
Assume $d \ge k$.
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
There exists universal constants $c_1,c_2$ such that
if the number of samples is
$n \ge \frac{c_1 k d^2 L R}{\epsilon}$
and $\lambda_n^{(2)} \ge \frac{c_2 L R}{\sqrt{n}}$,
and $\lambda_n^{(3)} \ge \frac{c_3 L R}{\sqrt{n}}$,
then the parameter estimates $\hat\theta$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \frac{\epsilon}{???}$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \epsilon ???$,
\end{theorem}
\begin{proof}
Let $T^* = M_3^*(W^*, W^*, W^*)$.
By Theorem 5.1 of \cite{AnandkumarGeHsu2012},
Need to convert error from $\epsilon$
hit by $(W^\top)^\dagger$
\end{proof}

>>>>

%%% second moments
Intuitively, performing regression on only $y$ given $x$ provides only first-order
information.  The key insight is that we can perform regression
on higher order powers to obtain more information about the parameters.
Specifically, for an integer $p \ge 1$, let us define the average
$p$-th order tensor power of the parameters as follows:
\begin{align}
M_p &\eqdef \sum_{h=1}^k \pi_h \beta_h\tp{p}. \label{eqn:Mp} % \in (\Re^{d})\tp{p}.
\end{align}
Now consider performing regression on $y^2$ given $x\tp{2}$.
Expanding $y^2 = (\innerp{\beta_h}{x} + \epsilon)^2$,
using the fact that $\innerp{\beta_h}{x}^p = \innerp{\beta_h\tp{p}}{x\tp{p}}$,
we have:
\begin{align}
y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h}{x} + (\epsilon^2 - \sigma^2). \nonumber
\end{align}
Again, we have expressed $y^2$ has a linear function of $x\tp{2}$
with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.\footnote{If $\sigma^2$ were not known,
we could treat it as another coefficient
to be estimated.  The coefficients $M_2$ and $\sigma^2$ can be estimated jointly
provided that $x$ does not already contain a bias ($x_j$ must be non-constant for every $j \in [d]$).}
Importantly, the noise has mean zero; 
in fact each of the three terms has mean zero
by (i) definition of $M_2$, (ii) independence of $\epsilon$ and $h$,
and (iii) the fact that $\E[\epsilon^2] = \sigma^2$.

@@

%%% third moment
Let us look to the third moment for additional information.
We can write $y^3$ as a linear function of $x\tp{3}$ with coefficients $M_3$,
a bias $3 \sigma^2 \innerp{M_1}{x}$ that can be estimated from \refeqn{y1}, and zero mean noise $\eta_3(x)$:
\begin{align}
y^3 &= \innerp{M_3}{x\tp{3}} + 3\sigma^2 \innerp{M_1}{x} + \eta_3(x), \label{eqn:y3} \\
\eta_3(x) &= \innerp{\beta_h\tp{3} - M_3}{x\tp{3}}
+ 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \nonumber \\
&\quad + 3(\epsilon^2 \innerp{\beta_h}{x} - \sigma^2 \innerp{M_1}{x})
+ \epsilon^3. \nonumber
\end{align}
Performing regression yields estimates for $M_3$.
It turns out that knowledge of $M_2$ and $M_3$ are sufficient to recover
all the parameters.

% Full algorithm
Now we are ready to state our full algorithm, which we call Spectral Experts
(\algorithmref{sec:algo}).
First, we perform three regressions to recover the compound parameters
$M_1$ \refeqn{y1},
$M_2$ \refeqn{y2}, and
$M_3$ \refeqn{y3}.
Since $M_2$ and $M_3$ both only have rank $k$,
we can use nuclear norm regularization
\cite{Tomioka2011,NegahbanWainwright2009}
to exploit this low-rank structure and improve our compound parameter estimates.
In the algorithm, the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$
are set to $\frac{1}{\sqrt{n}}$.
The resulting semidefinite program is a standard one which has received
much attention in recent years.
We use a rather simple gradient-based approach,
where the nuclear norm is handled in closed form by taking an SVD
and soft-thresholding the eigenvalues \cite{donoho95soft,cai10soft}.

% Tensor factorization
Having estimated the compound parameters $M_2$ and $M_3$,
it remains to recover the original parameters $\theta$.
\citet{AnandkumarGeHsu2012} showed that for $M_2$ and $M_3$ of
the forms in \refeqn{Mp}, it is possible to efficiently accomplish this.
Specifically, we first compute a whitening matrix $W$ based on the SVD of $M_2$
and use that to construct a tensor $T = M_3(W, W, W)$ whose factors are orthogonal.
We can use the robust tensor power method to compute all the
eigenvalues and eigenvectors of $T$, from which it is easy to recover
the parameters $\pi$ and $\{\beta_h\}$.

@@

In this section, we conduct a theoretical analysis of Spectral Experts.
In particular, we prove that it enjoys consistent parameter estimates.
\begin{theorem}[Convergence of Spectral Experts]
  \label{thm:convergence}
Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Assume that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for each $p \in \{1,2,3\}$.
Assume $d \ge k$.
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
There exists universal constants $c_1,c_2$ such that
if the number of samples is
$n \ge \frac{c_1 k d^2 L R}{\epsilon}$
and $\lambda_n^{(2)} \ge \frac{c_2 L R}{\sqrt{n}}$,
and $\lambda_n^{(3)} \ge \frac{c_3 L R}{\sqrt{n}}$,
then the parameter estimates $\hat\theta$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \frac{\epsilon}{???}$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \epsilon ???$,
\end{theorem}
\begin{proof}
Let $T^* = M_3^*(W^*, W^*, W^*)$.
By Theorem 5.1 of \cite{AnandkumarGeHsu2012},
Need to convert error from $\epsilon$
hit by $(W^\top)^\dagger$
\end{proof}

### 2013-02-15 doc/icml2013/intro.tex
theory done
Minor fixes

% Twist
In mixture of linear regressions, using third-order moments does not directly
reveal the tensor structure of the problem, so we cannot directly
apply the above tensor factorization techniques.
Our solution is to employ low-rank linear regression
\cite{NegahbanWainwright2009,Tomioka2011} on second
and third powers of the response to first recover the appropriate symmetric
tensors, from which we can then apply the tensor power method to retrieve the
final parameters.

@@

We use $x\tp{p}$ to represent the $p$-th order tensor
formed by taking the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots
i_p} = x_{i_1} \cdots x_{i_p}$. We will use
$\innerp{\cdot}{\cdot}$ to denote the generalized dot product between two $p$-th
order tensors: $\innerp{X}{Y} = \sum_{i_1, \ldots i_p} X_{i_1,
\ldots i_p} Y_{i_1, \ldots i_p}.$
A tensor $X$ is symmetric if for all $i,j \in [d]^p$ which are permutations,
$X_{i_1 \cdots i_p}$ = $X_{j_1 \cdots j_p}$ (all tensors in this paper will be symmetric).
For a $p$-th order tensor $X \in (\Re^d)\tp{p}$, let $X_{(i)} \in \Re^{d \times d^{p-1}}$
be the mode-$i$ unfolding of $X$.
%For $X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
%$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
%$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q}
%= \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
%\ldots, i_q}$.

@@

For a tensor $X \in (\Re^d)\tp{p}$,
let $\cvec(X) \in \Re^{\sum_{i=1}^p \binom{n}{i}}$ be
the collapsed vectorization of $X$: each component of $\cvec(X)$
is indexed by a vector of counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.
The value of that component is
$\sum_{k \in K(c)} X_{k_1 \cdots k_p}$,
where $K(c) = \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}| \}$ are the set of index vectors $k$ with that count profile.
For example, for $X \in \Re^{d \times d}$,
$\cvec(X) = (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.

>>>>

% Twist
In mixture of linear regressions, using third-order moments does not
directly reveal the tensor structure of the problem, so we cannot
directly apply the above tensor factorization techniques.  Our solution
is to employ low-rank linear regression
\cite{NegahbanWainwright2009,Tomioka2011} on second and third powers of
the response to recover the appropriate symmetric tensors, from which we
can then apply the tensor power method to retrieve the final parameters.

@@

We use $x\tp{p}$ to represent the $p$-th order tensor
formed by taking the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots
i_p} = x_{i_1} \cdots x_{i_p}$. We will use
$\innerp{\cdot}{\cdot}$ to denote the generalized dot product between two $p$-th
order tensors: $\innerp{X}{Y} = \sum_{i_1, \ldots i_p} X_{i_1,
\ldots i_p} Y_{i_1, \ldots i_p}.$
A tensor $X$ is symmetric if for all $i,j \in [d]^p$ which are permutations of each other,
$X_{i_1 \cdots i_p}$ = $X_{j_1 \cdots j_p}$ (all tensors in this paper will be symmetric).
For a $p$-th order tensor $X \in (\Re^d)\tp{p}$, let $X_{(i)} \in \Re^{d \times d^{p-1}}$
be the mode-$i$ unfolding of $X$.
%For $X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
%$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
%$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q}
%= \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
%\ldots, i_q}$.

@@

For a tensor $X \in (\Re^d)\tp{p}$,
let $\cvec(X) \in \Re^{\sum_{i=1}^p \binom{n}{i}}$ be
the collapsed vectorization of $X$: each component of $\cvec(X)$
is indexed by a vector of counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.
The value of that component is
$\sum_{k \in K(c)} X_{k_1 \cdots k_p}$,
where $K(c) = \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}| \}$ are the set of index vectors $k$ with that count profile.
For example, for $X \in \Re^{d \times d}$,
$\cvec(X) = (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
%For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.
